### 2024-07-19 chap_intro_ccl/conclusion.tex
First version of rebuttals completed + Grammarly Intro + Chap. 1 & 2

% WHAT WE DID (HOW?)
The general objective of this thesis has been to investigate the impact of uncertainties and limited foresight on the transition pathway of a whole-energy system, accounting for multi-energy carriers and multi-sectors, subject to \ce{CO2}-budget without prescribed trajectory for the reduction of the emissions.  We have studied the case study of Belgium with a focus on the role that renewable electrofuels (e-ammonia, e-methane, e-methanol and e-hydrogen) could play in this densely-populated and highly industrialised country with limited local renewable potentials. To have 66\% chance of limiting warming to 1.5°C and using the ``grandfathering'' sharing principle of the global budget, the \ce{CO2}-budget of the 2020-2050 transition of Belgium has been set to 1.2\,Gt$_{\ce{CO2},\text{eq}}$.

@@

The \acrfull{RL} approach was adopted to explore the myopic pathways respecting a \ce{CO2}-budget equivalent to 10 years of current emissions, without \ce{CO2}-trajectory. In this framework, an agent takes actions at the beginning of each time window to limit the emissions of the system, the \textit{environment}, and the consumption of fossil fuels. Starting from the initial state of the energy system in 2020, the agent takes every five years a set of actions until reaching 2050 where a transition was considered as successful when respecting the \ce{CO2}-budget. Repeating the whole transition with different sequences of actions-states allowed the agent coming up with an effective policy towards sustainability, considering the variation of the parameters of its environment.

@@

% THE RESULTS WE GOT (WHAT?)
Given the \ce{CO2}-budget to stick to a +1.5$^{\circ}$C increase of global temperature and the presumably high levels of demands,  aiming at a cost-optimum transition, Belgium should first opt for more efficiency to reduce by one third the primary energy consumption. Then comes the full deployment of local solar and wind energies that would represent about 30\% of the mix by 2050. Due to the limited potential of local renewables, electrofuels are found to be the third pillar of the Belgian transition. The uncertainty on the total cost of the transition is mainly due to the uncertainty on the cost of purchasing electrofuels (47\%), on the industrial end-use demand (23\%), on the discount rate (12\%) and on the price of fossil fuels (6\%).

@@

The Reinforcement Learning-based exploration of myopic transitions points out that near-term actions are mandatory to hope succeeding this ambitious transition.  Beyond 2030, if the right actions are not taken, the cumulative emissions are likely to overshoot the \ce{CO2}-budget. On top of strongly reducing the use of coal, fossil gas consumption should also be limited within 5 to 10 years. Once the fuel switch is operated from fossil gas to e-methane, putting a strong limit on the overall emissions is the most effective action to stick to the direction towards sustainability and encourage the import of renewable energy carriers. In the \gls{RL} framework, the agent's actions are mainly driven by the share of renewable energy carriers in the primary energy mix rather than the overall system efficiency. This share of renewable energy carriers shows intermediate milestones to reach, 62\%-share by 2030, to have higher chance to succeed the transition by 2050. Below this threshold, we find the so-called ``no-go zone'' where the chances of success are much lower. Due to limited foresight into the future and uncertainties like the possibility to install \gls{SMR} by 2040, successful myopic transitions rely more on importing renewable electrofuels than the perfect foresight approach. An illustrative example is the case of e-ammonia, for which imports by 2050 are in general twice larger in the myopic transition. This work also highlighted a major difference between the perfect foresight (PF) approach and the agent-based myopic transitions. Given the assumed full knowledge on the future, the trajectories to decrease the \ce{CO2} emissions resulting from perfect foresight are more progressive, and foreseen to be more applicable, compared to a sharper drop at the early stages of the transition, like in the myopic approach.

@@

Often brought up as a ``unicorn'' solution to the energy transition, nuclear \gls{SMR} would be a cost-competitive option against the import of electrofuels, especially e-ammonia and e-methane, but not against the deployment of local \acrfull{VRES} such as solar and wind. This confirmed that investing as soon as possible in local renewables is one pillar of the energy transition, especially when accounting for uncertainties and with the goal to respect an ambitious \ce{CO2}-budget of the transition.

@@

The Belgian transition pathways explored in this thesis massively relied on the early deployment of local \gls{VRES} and import of electrofuels. In some pathways, the current capacities of \gls{PV} and wind turbines are multiplied by 10 and 6 over a 5 to 10 year time window. This is much faster than the multiplication by 2 and 2.5 of PV and wind capacities from 2024 and 2034 projected by Elia, Belgium Transmission System Operator (TSO), in their adequacy analysis \cite{Elia_2024_2034}. The necessary amount of imported electrofuels could be 10 to 20 times higher than the potential that Belgium could benefit from considering the existing agreements with exporting countries \cite{lefebvre2022electrofuel}. A future work would be to constrain the deployment of local renewables or the availability of electrofuels. This could provide different pathways where electrofuels would be produced more locally using carbon capture technologies. This could also show the limits of the Belgian energy system to meet such an ambitious \ce{CO2}-budget in its energy transition when solely relying on technological progress, highlighting the need for demand reduction. One could then question the level of demands and address the question of the sufficiency. For a growing part of the scientific community, sufficiency is seen as a no-regret lever of action to reduce the anthropogenic \gls{GHG}. A study could consist in analysing the level of demands, within their respective sector, we should reach to stay within the \ce{CO2}-budget and limit the installation of solar PV, wind turbines, nuclear energy or the import of electrofuels.

@@

On the methodological side, several perspectives could be considered about the whole-energy system model or the implementation of \gls{RL}. The cumulative \acrfull{GHG} emissions of the transition has been constrained based on the use of the resources. However, in a system reducing drastically its emissions, indirect emissions such as the ones related to the construction and dismantling of assets would have a much higher share of the total emissions. Depending on the level of emission reduction, these indirect emissions could be one to three times higher than the direct emissions. To accurately integrate these emissions in the model and account for them in the \ce{CO2}-budget, \gls{LCA} can compute these indirect emissions. On top of this, \gls{LCA} could provide data to assess other impact indicators besides climate change \cite{astudillo2018integrating}. Based on this multi-criteria approach, another thread of future work would be to implement multi-objective optimisation \cite{dubois2023multi}. In this method, objectives other than the sole total transition cost such as the ones identified as planetary boundaries \cite{richardson2023earth} would be minimised.

>>>>

% WHAT WE DID (HOW?)
The general objective of this thesis has been to investigate the impact of uncertainties and limited foresight on the transition pathway of a whole-energy system, accounting for multi-energy carriers and multi-sectors, subject to \ce{CO2} budget without prescribed trajectory for the reduction of the emissions.  We have studied the case study of Belgium with a focus on the role that renewable electrofuels (e-ammonia, e-methane, e-methanol and e-hydrogen) could play in this densely-populated and highly industrialised country with limited local renewable potentials. To have 66\% chance of limiting warming to 1.5°C and using the ``grandfathering'' sharing principle of the global budget, the \ce{CO2} budget of the 2020-2050 transition of Belgium has been set to 1.2\,Gt$_{\ce{CO2},\text{eq}}$.

@@

The \acrfull{RL} approach was adopted to explore the myopic pathways respecting a \ce{CO2} budget equivalent to 10 years of current emissions, without \ce{CO2}-trajectory. In this framework, an agent takes actions at the beginning of each time window to limit the emissions of the system, the \textit{environment}, and the consumption of fossil fuels. Starting from the initial state of the energy system in 2020, the agent takes every five years a set of actions until reaching 2050 where a transition was considered as successful when respecting the \ce{CO2} budget. Repeating the whole transition with different sequences of actions-states allowed the agent coming up with an effective policy towards sustainability, considering the variation of the parameters of its environment.

@@

% THE RESULTS WE GOT (WHAT?)
Given the \ce{CO2} budget to stick to a +1.5$^{\circ}$C increase of global temperature and the presumably high levels of demands,  aiming at a cost-optimum transition, Belgium should first opt for more efficiency to reduce by one third the primary energy consumption. Then comes the full deployment of local solar and wind energies that would represent about 30\% of the mix by 2050. Due to the limited potential of local renewables, electrofuels are found to be the third pillar of the Belgian transition. The uncertainty on the total cost of the transition is mainly due to the uncertainty on the cost of purchasing electrofuels (47\%), on the industrial end-use demand (23\%), on the discount rate (12\%) and on the price of fossil fuels (6\%).

@@

The Reinforcement Learning-based exploration of myopic transitions points out that near-term actions are mandatory to hope succeeding this ambitious transition.  Beyond 2030, if the right actions are not taken, the cumulative emissions are likely to overshoot the \ce{CO2} budget. On top of strongly reducing the use of coal, fossil gas consumption should also be limited within 5 to 10 years. Once the fuel switch is operated from fossil gas to e-methane, putting a strong limit on the overall emissions is the most effective action to stick to the direction towards sustainability and encourage the import of renewable energy carriers. In the \gls{RL} framework, the agent's actions are mainly driven by the share of renewable energy carriers in the primary energy mix rather than the overall system efficiency. This share of renewable energy carriers shows intermediate milestones to reach, 62\%-share by 2030, to have higher chance to succeed the transition by 2050. Below this threshold, we find the so-called ``no-go zone'' where the chances of success are much lower. Due to limited foresight into the future and uncertainties like the possibility to install \gls{SMR} by 2040, successful myopic transitions rely more on importing renewable electrofuels than the perfect foresight approach. An illustrative example is the case of e-ammonia, for which imports by 2050 are in general twice larger in the myopic transition. This work also highlighted a major difference between the perfect foresight (PF) approach and the agent-based myopic transitions. Given the assumed full knowledge on the future, the trajectories to decrease the \ce{CO2} emissions resulting from perfect foresight are more progressive, and foreseen to be more applicable, compared to a sharper drop at the early stages of the transition, like in the myopic approach.

@@

Often brought up as a ``unicorn'' solution to the energy transition, nuclear \gls{SMR} would be a cost-competitive option against the import of electrofuels, especially e-ammonia and e-methane, but not against the deployment of local \acrfull{VRES} such as solar and wind. This confirmed that investing as soon as possible in local renewables is one pillar of the energy transition, especially when accounting for uncertainties and with the goal to respect an ambitious \ce{CO2} budget of the transition.

@@

The Belgian transition pathways explored in this thesis massively relied on the early deployment of local \gls{VRES} and import of electrofuels. In some pathways, the current capacities of \gls{PV} and wind turbines are multiplied by 10 and 6 over a 5 to 10 year time window. This is much faster than the multiplication by 2 and 2.5 of PV and wind capacities from 2024 and 2034 projected by Elia, Belgium Transmission System Operator (TSO), in their adequacy analysis \cite{Elia_2024_2034}. The necessary amount of imported electrofuels could be 10 to 20 times higher than the potential that Belgium could benefit from considering the existing agreements with exporting countries \cite{lefebvre2022electrofuel}. A future work would be to constrain the deployment of local renewables or the availability of electrofuels. This could provide different pathways where electrofuels would be produced more locally using carbon capture technologies. This could also show the limits of the Belgian energy system to meet such an ambitious \ce{CO2} budget in its energy transition when solely relying on technological progress, highlighting the need for demand reduction. One could then question the level of demands and address the question of the sufficiency. For a growing part of the scientific community, sufficiency is seen as a no-regret lever of action to reduce the anthropogenic \gls{GHG}. A study could consist in analysing the level of demands, within their respective sector, we should reach to stay within the \ce{CO2} budget and limit the installation of solar PV, wind turbines, nuclear energy or the import of electrofuels.

@@

On the methodological side, several perspectives could be considered about the whole-energy system model or the implementation of \gls{RL}. The cumulative \acrfull{GHG} emissions of the transition has been constrained based on the use of the resources. However, in a system reducing drastically its emissions, indirect emissions such as the ones related to the construction and dismantling of assets would have a much higher share of the total emissions. Depending on the level of emission reduction, these indirect emissions could be one to three times higher than the direct emissions. To accurately integrate these emissions in the model and account for them in the \ce{CO2} budget, \gls{LCA} can compute these indirect emissions. On top of this, \gls{LCA} could provide data to assess other impact indicators besides climate change \cite{astudillo2018integrating}. Based on this multi-criteria approach, another thread of future work would be to implement multi-objective optimisation \cite{dubois2023multi}. In this method, objectives other than the sole total transition cost such as the ones identified as planetary boundaries \cite{richardson2023earth} would be minimised.
